---
title: "bonobo brain"
author: "Josema"
date: "2025-01-27"
output: html_document
---

# bonobo

# multiome analysis de human

```{r}
library(tidyverse)
library(Signac)
library(Seurat)

options(timeout = 600)

library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(SeuratObject)
```


```{r}
counts <- Read10X_h5("/users/genomics/josema/multiome/Bonobo_humanindex/outs/filtered_feature_bc_matrix.h5")
fragpath <- "/users/genomics/josema/multiome/Bonobo_humanindex/outs/atac_fragments.tsv.gz"
```

```{r}
# get gene annotations for hg38
annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevels(annotation) <- paste0('chr', seqlevels(annotation))

# create a Seurat object containing the RNA adata
bonobo_unfiltered <- CreateSeuratObject(
  counts = counts$`Gene Expression`,
  assay = "RNA"
)

# create ATAC assay and add it to the object
bonobo_unfiltered[["ATAC"]] <- CreateChromatinAssay(
  counts = counts$Peaks,
  sep = c(":", "-"),
  fragments = fragpath,
  annotation = annotation
)
```

## atac quality control

```{r}
DefaultAssay(bonobo_unfiltered) <- "ATAC"

bonobo_unfiltered <- NucleosomeSignal(bonobo_unfiltered)
bonobo_unfiltered <- TSSEnrichment(bonobo_unfiltered)
```


```{r fig.height=4, fig.width=4}
pdf('/users/genomics/josema/quality_atac_gex_bonobo_humanindexed_clustered.pdf')
DensityScatter(bonobo_unfiltered, x = 'nCount_ATAC', y = 'TSS.enrichment', log_x = TRUE, quantiles = TRUE)
features=c("nCount_RNA",
           "nCount_ATAC",
           "TSS.enrichment",
           "nucleosome_signal",
           'Nuclear_fraction',
           'percent.mt',
           'nFeature_RNA'
           )
```


```{r}
VlnPlot(
  object = bonobo_unfiltered,
  features = features,
  ncol = 4,
  pt.size = 0,
  cols = colors
)+
  geom_hline(yintercept = 100, color = "gray20") 
```


```{r fig.height=4, fig.width=4}
colors=c('1'='#48cae4', '2'='#a4d5a6','3'='#80c684',
  '4'='#68bb6c','5'='#ffad42','6'='#0077b6',
  '7'='#ef476f','0'='#0096c7' )

DimPlot(object = bonobo_unfiltered, reduction = "umap",  cols = rep("grey",8))+
  theme(aspect.ratio = 1)

DimPlot(object = bonobo_unfiltered, reduction = "umap",  cols = colors)+
  theme(aspect.ratio = 1)+
  theme_void()+
  guides(color="none")

DimPlot(object = bonobo_unfiltered, reduction = "umap.atac")


bonobo_unfiltered[["percent.mt"]] <- PercentageFeatureSet(bonobo_unfiltered, pattern = "^MT-")

FeaturePlot(bonobo_unfiltered, features=features, reduction = 'umap')

FeaturePlot(bonobo_unfiltered, features='percent.mt' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('Mitochontrial genes percentage')+
  scale_color_gradient(low = 'gray95', high = '#184e77') 

FeaturePlot(bonobo_unfiltered, features='MALAT1' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('MALAT1 Expression')+
  scale_color_gradient(low = 'gray95', high = '#adc178')

FeaturePlot(bonobo_unfiltered, features='nFeature_RNA' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('Nº of RNA features')+
  scale_color_gradient(low = 'gray95', high = '#ef233c')

  

FeatureScatter(bonobo_unfiltered, feature1 = "percent.mt", feature2 = "nFeature_RNA")+ scalecolorman
dev.off()

```

```{r}
saveRDS(bonobo_unfiltered, file = '/users/genomics/josema/multiome/rmd_files/bonobo_unfiltered.rmd')


bonobo_unfiltered <- readRDS('/users/genomics/josema/multiome/rmd_files/bonobo_unfiltered.rmd')
```
v
```{r}
characteristics <- read.csv('/users/genomics/josema/multiome/Bonobo_humanindex/outs/per_barcode_metrics.csv')

mapped_metadata <- characteristics %>%  
  dplyr::filter(is_cell==1) %>% 
  dplyr::select(barcode,gex_mapped_reads, gex_conf_exonic_reads,gex_conf_intronic_reads) %>% 
  column_to_rownames('barcode') %>% 
  dplyr::mutate(Nuclear_fraction=gex_conf_intronic_reads/(gex_conf_intronic_reads+gex_conf_exonic_reads)) 
```

```{r}
bonobo_unfiltered@meta.data <-  cbind(bonobo_unfiltered@meta.data,mapped_metadata[rownames(bonobo_unfiltered@meta.data),]) 
DefaultAssay(bonobo_unfiltered) <- "RNA"
bonobo_unfiltered[["percent.mt"]] <- PercentageFeatureSet(bonobo_unfiltered, pattern = "^MT-")

```



```{r}
DefaultAssay(bonobo_unfiltered) <- "RNA"
bonobo_unfiltered <- NormalizeData(bonobo_unfiltered)
bonobo_unfiltered <- FindVariableFeatures(bonobo_unfiltered)
bonobo_unfiltered <- ScaleData(bonobo_unfiltered, , features=rownames(bonobo_filtered))
  bonobo_unfiltered <- RunPCA(bonobo_unfiltered, 
                              features= VariableFeatures(bonobo_unfiltered))
bonobo_unfiltered <- FindNeighbors(bonobo_unfiltered, dims = 1:10)
bonobo_unfiltered <- FindClusters(bonobo_unfiltered, resolution = 0.5)
bonobo_unfiltered <- RunUMAP(bonobo_unfiltered, dims = 1:30)

DefaultAssay(bonobo_unfiltered) <- "ATAC"
bonobo_unfiltered <- RunTFIDF(bonobo_unfiltered)
bonobo_unfiltered <- FindTopFeatures(bonobo_unfiltered, min.cutoff = "q0")
bonobo_unfiltered <- RunSVD(bonobo_unfiltered)
# We exclude the first dimension as this is typically correlated with sequencing depth
bonobo_unfiltered <- RunUMAP(bonobo_unfiltered, reduction = "lsi", dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
```




## clustering

```{r}
BiocManager::install("celldex")
BiocManager::install("SingleR")

library(celldex)
library(SingleR)
```

```{r}
ref <- celldex::HumanPrimaryCellAtlasData()
unique(ref$label.main)
unique(ref$label.fine)


ref <- celldex::HumanPrimaryCellAtlasData()
unique(ref$label.main)
unique(ref$label.fine)
```

```{r}
bonobo_unfiltered %>%  write
```


# Local analysis

```{r}
library(Seurat)
library(Signac)
library(tidyverse)
bonobo_unfiltered <- readRDS('~/Downloads/bonobo_unfiltered.rmd')
bonobo_unfiltered@assays$RNA$counts
```


## Marcadores de tipos celulares
```{r fig.height=40, fig.width=20}
markers_neurodev <-  c('STMN2','MAP2','DCX','GAP43','SYT5', #Neuron
                 'TBR1','SLC17A7','SLC17A6', # EN
                 'SLC32A1','GAD2','DLX2','ASCL1', # IN
                 'NHLH1','NEUROD4', # nEN
                 'EOMES','NEUROG1', # IPC
                 'HOPX','MOXD1','PTN', #oRG
                 'HES1','PAX6','SOX2','ID4',#RG
                 'SAMD4A','CCN2', #tRG
                 'ASPM', # cell div
                 'FOXJ1','GMNC' #Epend
                 )

adult_markers <-  c('INA', 
                    'SLC17A6', 'SLC17A7',
                    'SLC32A1',
                    'PTPRC',
                    'CLDN5', 'ACTA2',
                    'LUM', 
                    'PDGFRA', 'SOX10',
                    'PLP1',
                    'AQP4',
                    'FOXJ1',
                    'TTR')

adult_brain_cortex_markers <- c(
  # Neurons
  "RBFOX3", "MAP2", "SYT1", "SNAP25", "NCAM1", "DCX", "STMN2", "BCL11B",
  
  # Astrocytes
  "GFAP", "AQP4", "S100B", "ALDH1L1", "SLC1A2", "SOX9", "CLDN10", "CSPG4", "TNC",
  
  # Oligodendrocytes
  "MBP", "PLP1", "MOG", "MAG", "CNP", "SOX10", "CLDN11", "MOBP", "NKX2-2",
  
  # Microglia
  "CD68", "AIF1", "TMEM119", "CX3CR1", "P2RY12", "CSF1R", "CD86", "ITGAM",
  
  # Excitatory Neurons
  "SLC17A7", "CAMK2A", "GRIN1", "NRGN", "FOXP1", "SATB2", "KCNQ5",
  
  # Inhibitory Neurons
  "GAD1", "GAD2", "DLX1", "DLX2", "SST", "PVALB", "VIP", "NPY", "LHX6", "ARX",
  
  # Oligodendrocyte Progenitor Cells
  "OLIG2", "PDGFRA", "ASCL1", "CD44", "FABP7", "TCF7L2",
  
  # Upper Layer Cortical Neurons
  "CUX1", "RORB", "LAMP5", "FEZF1", "FOXP2", "TLE4",
  
  # Deep Layer Cortical Neurons
  "TBR1", "FEZF2", "SOX5", "MYT1L", "EBF1",
  
  # Mesenchymal Cells
  "VIM", "ACTA2", "PDGFRB", "COL1A1", "THY1", "FAP", "ALDH1A1", "ITGA5"
)



DefaultAssay(bonobo_unfiltered) <- "RNA"

FeaturePlot(bonobo_unfiltered, features=markers_neurodev , reduction = 'umap', order = T)
FeaturePlot(bonobo_unfiltered, features=markers_neurodev , reduction = 'umap', order = T)

FeaturePlot(bonobo_unfiltered, features=adult_markers , reduction = 'umap', order = T)
FeaturePlot(bonobo_unfiltered, features=adult_brain_cortex_markers , reduction = 'umap', order = T)


```


RBFOX3:

    Appeared in the Neurons section twice.

BCL11B:

    Appeared in:
        Neurons section
        Excitatory Neurons section
        Deep Layer Cortical Neurons section

SOX10:

    Appeared in:
        Oligodendrocytes section
        Oligodendrocyte Progenitor Cells section

SATB2:

    Appeared in:
        Excitatory Neurons section
        Upper Layer Cortical Neurons section

CSPG4:

    Appeared in:
        Astrocytes section
        Oligodendrocyte Progenitor Cells section

CD44:

    Appeared in:
        Oligodendrocyte Progenitor Cells section
        Mesenchymal Cells section

TLE4:

    Appeared in:
        Upper Layer Cortical Neurons section
        Deep Layer Cortical Neurons section
        

```{r }
bonobo_unfiltered_pseudobulk <- AggregateExpression(bonobo_unfiltered, assays = 'RNA', return.seurat = T, group.by ='seurat_clusters')

DimPlot(bonobo_unfiltered, reduction = 'umap', label = T)+
  theme(aspect.ratio = 1)
DimPlot(bonobo_unfiltered, reduction = 'umap.atac' )+
  theme(aspect.ratio = 1)


DoHeatmap(bonobo_unfiltered_pseudobulk, features = adult_brain_cortex_markers) +
   scale_fill_gradientn(colors = c("blue", "white", "red"))
```


```{r fig.height=3.5, fig.width=15}
library(ComplexHeatmap)

data_heatmap <- bonobo_unfiltered_pseudobulk@assays$RNA$scale.data %>%
  t() %>% 
  as.data.frame() %>% 
  select(adult_brain_cortex_markers) 

levels=c("Neurons", "Astrocytes", "Oligodendrocytes", "Microglia", 
              "Excitatory Neurons", "Inhibitory Neurons", 
              "OPC", "Upper Layer CN", 
              "Deep Layer CN", "Mesenchymal Cells")

data_heatmap%>% 
  as.data.frame() %>% 
  Heatmap(cluster_columns  = F,
          column_order = colnames(data_heatmap),
          column_split =  factor(rep(levels ,c(8,9,9,8,7,10,6,6,5,8)),
                               levels=levels),
          column_title_gp = gpar(fontsize = 10)
         )
          

```




```{r fig.height=4, fig.width=4}
# Create a new metadata column named 'QC_filter' with 'Good' and 'Bad' labels
bonobo_unfiltered$QC_filter <- ifelse(bonobo_unfiltered$nFeature_RNA > 100, "Good", "Bad")

# Plot with DimPlot, coloring based on QC_filter
DimPlot(bonobo_unfiltered, group.by = c("QC_filter")) +
  scale_color_manual(values = c("Bad" = "gray90", "Good" = "#80c684"))+
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)

```


```{r fig.height=4, fig.width=4}
umap_df <- as.data.frame(Embeddings(bonobo_unfiltered, "umap"))
umap_df$cluster <- bonobo_unfiltered$seurat_clusters
umap_df$QC_filter <- bonobo_unfiltered$QC_filter

# Define transparency levels
umap_df$alpha_value <- ifelse(umap_df$QC_filter == "Good", 1, 0.05)

# Create UMAP plot with cluster colors and QC-based alpha
ggplot(umap_df, aes(x = umap_1, y = umap_2, color = as.factor(cluster), alpha = alpha_value)) +
  geom_point(show.legend = F, size=0.8) +
  scale_alpha_identity() +  # Ensures correct alpha values
  theme_void() +
  scale_color_manual(values = colors)+
  theme(aspect.ratio = 1)

```

# Bonobo filtered

```{r}
bonobo_filtered <- subset(
  x = bonobo_unfiltered,
  subset = nFeature_RNA  > 100)

DefaultAssay(bonobo_filtered) <- "RNA"
bonobo_filtered <- NormalizeData(bonobo_filtered)
bonobo_filtered <- FindVariableFeatures(bonobo_filtered, nfeatures = 7000)
bonobo_filtered <- ScaleData(bonobo_filtered,features=rownames(bonobo_filtered))
bonobo_filtered <- RunPCA(bonobo_filtered, 
                              features= VariableFeatures(bonobo_filtered))
bonobo_filtered <- FindNeighbors(bonobo_filtered, dims = 1:10, )
bonobo_filtered <- FindClusters(bonobo_filtered, resolution = 0.5)
bonobo_filtered <- RunUMAP(bonobo_filtered, dims = 1:30)

DefaultAssay(bonobo_filtered) <- "ATAC"
bonobo_filtered <- RunTFIDF(bonobo_filtered)
bonobo_filtered <- FindTopFeatures(bonobo_filtered, min.cutoff = "q0")
bonobo_filtered <- RunSVD(bonobo_filtered)
# We exclude the first dimension as this is typically correlated with sequencing depth
bonobo_filtered <- RunUMAP(bonobo_filtered, reduction = "lsi", dims = 2:30, reduction.name = "umap.atac", reduction.key = "atacUMAP_")
```

```{r}
bonobo_multimodal <-
  FindMultiModalNeighbors(
  object = bonobo_filtered,
  reduction.list = list("pca", "lsi"),
  dims.list = list(1:20, 2:10),
  modality.weight.name = "RNA.weight",
  verbose = TRUE
)

bonobo_multimodal <- RunUMAP(
  object = bonobo_multimodal,
  nn.name = "weighted.nn",
  assay = "RNA",
  verbose = TRUE
)
DefaultAssay(bonobo_multimodal) <- 'RNA'

bonobo_multimodal <- FindClusters(
  object = bonobo_multimodal,
  graph.name = "wsnn",  # weighted shared nearest neighbor graph created by FindMultiModalNeighbors
  resolution = 0.8,  # Adjust resolution. 0.8 is the default
)



```

# link peaks to genes

```{r fig.height=4, fig.width=4}
# dimplot del RNA

bonobo_filtered[["percent.mt"]] <- PercentageFeatureSet(bonobo_filtered, pattern = "^MT-")


FeaturePlot(bonobo_filtered, features='percent.mt' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('Mitochontrial genes percentage')+
  scale_color_gradient(low = 'gray95', high = '#184e77') 

FeaturePlot(bonobo_filtered, features='MALAT1' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('MALAT1 Expression')+
  scale_color_gradient(low = 'gray95', high = '#adc178')

FeaturePlot(bonobo_filtered, features='nFeature_RNA' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('N of RNA features')+
  scale_color_gradient(low = 'gray95', high = '#ef233c')

DimPlot(bonobo_filtered, reduction = 'umap', cols = colors) +
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)

FeaturePlot(bonobo_filtered,
            reduction = 'umap.atac',
            features = c('nucleosome_signal'))+
  theme(aspect.ratio = 1)+
  ggtitle('Nucleosome signal')+
  scale_color_gradient(low = 'gray95', high = '#184e77') 

FeaturePlot(bonobo_filtered,
            reduction = 'umap.atac',
            features = c('TSS.enrichment'))+
  theme(aspect.ratio = 1)+
  ggtitle('TSS enrichment')+
  scale_color_gradient(low = 'gray95', high = '#adc178') 

FeaturePlot(bonobo_filtered,
            reduction = 'umap.atac',
            features = c('nFeature_ATAC'))+
  theme(aspect.ratio = 1)+
  ggtitle('Number of Features')+
  scale_color_gradient(low = 'gray95', high = '#ef233c') 


# Dimplot del ATAC

DimPlot(bonobo_filtered, reduction = 'umap.atac' ,cols = colors) +
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)



# Dimplot del multimodal
DimPlot(bonobo_multimodal, cols = colors) +
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)
```

save bonobo filtered and multimodal

```{r}
saveRDS(bonobo_filtered, '/home/josema/Downloads/bonobo_filtered.rds')
saveRDS(bonobo_multimodal, '/home/josema/Downloads/bonobo_muktimodal.rds')


bonobo_filtered <- readRDS('/home/josema/Downloads/bonobo_filtered.rds')
```


# heatmap multimodal 
```{r}
bonobo_multimodal_pseudobulk <- AggregateExpression(bonobo_multimodal, assays = 'RNA', return.seurat = T, group.by ='seurat_clusters')
```


```{r fig.height=3.5, fig.width=15}
data_heatmap_multimodal <- bonobo_multimodal_pseudobulk@assays$RNA$scale.data %>%
  t() %>% 
  as.data.frame() %>% 
  select(adult_brain_cortex_markers) 

levels=c("Neurons", "Astrocytes", "Oligodendrocytes", "Microglia", 
              "Excitatory Neurons", "Inhibitory Neurons", 
              "OPC", "Upper Layer CN", 
              "Deep Layer CN", "Mesenchymal Cells")

data_heatmap_multimodal%>% 
  as.data.frame() %>% 
  Heatmap(cluster_columns  = F,
          column_order = colnames(data_heatmap_multimodal),
          column_split =  factor(rep(levels ,c(8,9,9,8,7,10,6,6,5,8)),
                               levels=levels),
          column_title_gp = gpar(fontsize = 10)
         )
```

# heatmap filtered

```{r}
bonobo_filtered_pseudobulk <- AggregateExpression(bonobo_filtered, assays = 'RNA', return.seurat = T, group.by ='seurat_clusters')
```


```{r fig.height=3, fig.width=15}

devtools::install_github("dgrtwo/gganimate")

data_heatmap_filtered <- bonobo_filtered_pseudobulk@assays$RNA$scale.data %>%
  t() %>% 
  as.data.frame() %>% 
  select(adult_brain_cortex_markers) 

levels=c("Neurons", "Astrocytes", "Oligodendrocytes", "Microglia", 
              "Excitatory Neurons", "Inhibitory Neurons", 
              "OPC", "Upper Layer CN", 
              "Deep Layer CN", "Mesenchymal Cells")

data_heatmap_filtered%>% 
  as.data.frame() %>% 
  Heatmap(cluster_columns  = F,
          column_order = colnames(data_heatmap_filtered),
          column_split =  factor(rep(levels ,c(8,9,9,8,7,10,6,6,5,8)),
                               levels=levels),
          column_title_gp = gpar(fontsize = 10)
         )
```

# UMAP with different

```{r}
bonobo_filtered_2 <- bonobo_filtered



DefaultAssay(bonobo_filtered_2) <- "RNA"
bonobo_filtered_2 <- FindNeighbors(bonobo_filtered_2,
                                   dims = 1:10,
                                   k.param = 5) 
bonobo_filtered_2 <- FindClusters(bonobo_filtered_2,
                                  resolution = 0.8) # above 1, higher number of communities
bonobo_filtered_2 <- RunUMAP(bonobo_filtered_2,
                             dims = 1:5,
                             min.dist = 0.3,  # Default is 0.3; lower keeps more structure
                             n.neighbors = 5,  # Default is 30; lower values keep more local structure
                             )
DimPlot(bonobo_filtered_2, reduction = 'umap') +
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)

```

```{r}

FeaturePlot(bonobo_filtered_2, features='percent.mt' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('Mitochontrial genes percentage')+
  scale_color_gradient(low = 'gray95', high = '#184e77') 

FeaturePlot(bonobo_filtered_2, features='MALAT1' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('MALAT1 Expression')+
  scale_color_gradient(low = 'gray95', high = '#adc178')

FeaturePlot(bonobo_filtered_2, features='nFeature_RNA' , reduction = 'umap', order = T)+
  theme(aspect.ratio = 1)+
  ggtitle('N of RNA features')+
  scale_color_gradient(low = 'gray95', high = '#ef233c')

DimPlot(bonobo_filtered_2, reduction = 'umap') +
    theme_void()+
  theme(aspect.ratio = 1,
        legend.position = NULL)+
    labs(title = NULL)

ElbowPlot(bonobo_filtered)
```



```{r fig.height=20, fig.width=20}
library(ggpubr)  # For arranging plots compactly

test_umap_parameters <- function(seurat_obj) {
  # Define parameter ranges
  k_values <- c(20, 40, 60)
  resolution_values <- c(0.4, 0.8, 1.2, 1.6)
  min_dist_values <- c(0.1, 0.3, 0.5)
  n_neighbors_values <- c(30, 50, 70)
  
  plot_list <- list()
  count <- 1  # Track plot index
  
  # Loop through all combinations
  for (k in k_values) {
    for (res in resolution_values) {
      for (min_d in min_dist_values) {
        for (n_neigh in n_neighbors_values) {
          
          # Clone original object to avoid modifying input
          temp_obj <- seurat_obj
          
          # Run Seurat steps with different parameters
          temp_obj <- FindNeighbors(temp_obj, dims = 1:30, k.param = k)
          temp_obj <- FindClusters(temp_obj, resolution = res)
          temp_obj <- RunUMAP(temp_obj, dims = 1:30, min.dist = min_d, n.neighbors = n_neigh)
          
          # Create UMAP plot with parameters in subtitle
          p <- DimPlot(temp_obj, reduction = 'umap') +
            theme_void() +
            theme(aspect.ratio = 1, legend.position = "none", plot.margin = margin(1, 1, 1, 1, "mm")) +
            labs(subtitle = paste0("k=", k, ", res=", res, ",\n min.dist=", min_d, ", n.neigh=", n_neigh))
          
          # Store plot in list
          plot_list[[count]] <- p
          count <- count + 1
        }
      }
    }
  }
  
  # Arrange all plots compactly using ggarrange
  combined_plot <- ggarrange(plotlist = plot_list, ncol = 4, nrow = 5)  
  
  return(combined_plot)
}

# Run the function on your dataset
plots <- test_umap_parameters(bonobo_filtered_2)

pdf('/home/josema/Downloads/plots_variables.pdf', width = 20, height = 20)
print(plots)  # Ensure the plot is printed inside the PDF
dev.off()
```

# create a gif 

```{r}
library(Seurat)
library(ggplot2)

BiocManager::install('udunits2')
library(gganimate)
library(dplyr)

# Function to test different UMAP parameters and store results
generate_umap_data <- function(seurat_obj) {
  # Define parameter ranges
  k_values <- c(20, 40, 60)
  resolution_values <- c(0.4, 0.8, 1.2, 1.6)
  min_dist_values <- c(0.1, 0.3, 0.5)
  n_neighbors_values <- c(5, 20, 70)

  umap_data_list <- list()
  frame_count <- 1
  
  # Loop through all parameter combinations
  for (k in k_values) {
    for (res in resolution_values) {
      for (min_d in min_dist_values) {
        for (n_neigh in n_neighbors_values) {
          
          # Clone original object to avoid modifying input
          temp_obj <- seurat_obj
          
          # Run Seurat steps with different parameters
          temp_obj <- FindNeighbors(temp_obj, dims = 1:30, k.param = k, verbose = FALSE)
          temp_obj <- FindClusters(temp_obj, resolution = res, verbose = FALSE)
          temp_obj <- RunUMAP(temp_obj, dims = 1:30, min.dist = min_d, n.neighbors = n_neigh, verbose = FALSE)

          # Extract UMAP coordinates
          umap_df <- as.data.frame(Embeddings(temp_obj, "umap"))
          umap_df$Cluster <- as.factor(Idents(temp_obj))  # Add cluster identity
          umap_df$Frame <- frame_count  # Frame number for animation
          umap_df$Params <- paste0("k=", k, ", res=", res, ", min.dist=", min_d, ", n.neigh=", n_neigh)

          umap_data_list[[frame_count]] <- umap_df
          frame_count <- frame_count + 1
        }
      }
    }
  }

  # Combine all UMAP data
  umap_data <- bind_rows(umap_data_list)
  return(umap_data)
}

# Function to create animated UMAP plot
create_umap_animation <- function(umap_data, gif_path = "umap_animation.gif") {
  p <- ggplot(umap_data, aes(x = UMAP_1, y = UMAP_2, color = Cluster)) +
    geom_point(alpha = 0.7, size = 1.5) +  # Adjust point size and transparency
    theme_minimal(base_size = 14) +  # Use minimal theme for clarity
    theme(legend.position = "none") +
    labs(title = "UMAP Parameter Testing", subtitle = "{closest_state}") +  
    transition_states(Params, transition_length = 2, state_length = 1) +
    ease_aes('linear')

  # Save as GIF
  anim_save(gif_path, animation = p, width = 8, height = 6, units = "in", res = 150)
  message("GIF saved at: ", gif_path)
}

# Run the pipeline
umap_data <- generate_umap_data(bonobo_filtered_2)
create_umap_animation(umap_data)


```


# mapping to kanton dataset

```{r}
library(tidyverse)
kanton <- readRDS('/home/josema/Downloads/org2m_human_singlecells_GRCh38.rds')

kanton <- FindVariableFeatures(kanton, verbose = FALSE)
kanton <- RunPCA(kanton, verbose = FALSE)

raw_counts <- Matrix::readMM(file = '/home/josema/Downloads/human_cell_counts_consensus.mtx.gz') %>%  Matrix::Matrix(data=., sparse=T)




anchors <- FindTransferAnchors(
  reference = kanton,
  query = bon,
  normalization.method = "SCT",  # Use "SCT" if using SCTransform; otherwise, use "LogNormalize"
  reference.reduction = "pca",
  dims = 1:30
)


VlnPlot(object = kanton$celltype_reannot, features = 'NEUROG2')


Seurat::vio
```

