---
title: "Cannabinoid receptor"
output: html_document
date: "2024-04-08"
---

# Introduction

There are two main cannabinoid receptors: CB1 and CB2. This receptors are encoded by CNR1 and CNR2 genes, located in Chr6 and Chr1.

The role of CNR1 in development and the consecuences that exogenous factors (drugs) might have in brain and CNS organogenesis.

We usually use mice as experimental models to study the role of drug intake during pregnancy, but there are evidences that the cannabinoid receptor transcriptional pattern of mouse and human might be different.

*Is the pattern of cannabinoid receptor expression along neurogenesis similar in mouse and human?*

# Objectives

-   Find mouse and human databases

# Analysis using NOAK's dataset

```{r}
library(tidyverse)
library(Seurat)
library(readr) 
#library(SeuratDisk)
```

```{r}
# Load the data
expression<-readRDS("~/Downloads/merged_scRNA_unfiltered_IDs_faye.RDS")
```

```{r}
# create the metadata

#para la expresion, cargo el data. aunq tambien esta el scaled.data. no se 100% bien cual de los dos deberia cargar. no se por que han escalado los datos

expression_matrix<-expression@assays$RNA@data
metadata<-expression@meta.data
metadata$celltype<-Idents(expression)
#mezclar las mitoticas con las no mitoticas.
metadata$celltype<-gsub("_M","",metadata$celltype)
metadata<-metadata %>% filter(!is.na(pseudotime))
table(metadata$celltype)/175
#hacer pseudobulks de 175 celulas. pq es lo mas justo para hacer 2 pseudobulks en pn3

#hacer los heatmaps con esto. y luego ponerle los picos, su trayectoria, motivos, y spacing. y
j<-1
for (cell in c("NSC","IPC","PN1","PN2","PN3")) {
  print(cell)
  metadata_cell<-metadata %>% filter(celltype==cell)
  metadata_cell<-metadata_cell %>% arrange(pseudotime)
  i<-1
  while ((i+175)<=length(metadata_cell$pseudotime)) {
    pseudobulk_cells<-rownames(metadata_cell)[i:(i+174)]
    pseudobulk_expression<-rowSums(as.matrix(expression_matrix[,pseudobulk_cells]))
    if (j==1) { pseudo_expression<-pseudobulk_expression}
    if (j>1) { pseudo_expression<-cbind(pseudo_expression,pseudobulk_expression)}
    i<-i+175
    j<-j+1
  }
}
colnames(pseudo_expression)
colnames(pseudo_expression)<-c(paste0("NSC_",1:11),paste0("IPC_",1:8),paste0("PN1_",1:9),paste0("PN2_",1:8),paste0("PN3_",1:2))
```

```{r}
z_score <- pseudo_expression %>% 
  as.data.frame() %>% 
  rownames_to_column('Symbol') %>%
  pivot_longer(-Symbol, names_to = 'Cell_type', values_to = 'expression') %>% 
  group_by(Symbol) %>% 
  mutate(Zscore=(expression - mean(expression))/sd(expression)) %>%
  ungroup() %>% 
  mutate(Group=gsub('\\_.*','',Cell_type),
         Group=factor(Group, levels=c('NSC','IPC','PN1','PN2','PN3')),
         Pseudotime=gsub('.*\\_','',Cell_type),
         Pseudotime= factor(Pseudotime, levels = as.character(1:12)),
         Symbol=toupper(Symbol))
```

```{r, fig.width=6, fig.height=1.6}

z_score %>% 
  filter(grepl('CNR[[:digit:]]',Symbol)) %>% 
  ungroup() %>% 
  ggplot(aes(x=Pseudotime, y=Symbol, fill= expression))+
  scale_fill_gradient(low = "#caf0f8",high = '#03045e') +
  geom_tile()+
  facet_grid(~Group, scales  = 'free', space = 'free')+
  theme_bw()+
  theme(legend.position = 'bottom',
        strip.placement = "outside", 
        strip.background.x=element_rect(color = NA,
                                        fill=NA), 
        strip.text.x.top = element_text(face = 'bold', color = 'gray30'),
        axis.title.y=element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x=element_blank())


z_score %>% 
  filter(grepl('CNR[[:digit:]]',Symbol)) %>% 
  ungroup() %>% 
  ggplot(aes(x=Pseudotime, y=Symbol, fill= Zscore))+
  scale_fill_gradient2(  low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red")) +
  geom_tile()+
  facet_grid(~Group, scales  = 'free', space = 'free')+
  theme_bw()+
  theme(legend.position = 'bottom',
        strip.placement = "outside", 
        strip.background.x=element_rect(color = NA,
                                        fill=NA), 
        strip.text.x.top = element_text(face = 'bold', color = 'gray30'),
        axis.title.y=element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x=element_blank())
```

# La mano 2021 data

```{r}
Manno_mouse <- readRDS('/projects_ng/SC_DATA/Manno2021/mdev_seu_0503_2021.rds')
Manno_mouse <- UpdateSeuratObject(Manno_mouse)
```

```{r}
Manno_metadata <-  Manno_mouse@meta.data
Manno_RNA <- Manno_mouse[["RNA"]]$data %>% Matrix::Matrix(as.matrix(.), sparse = T)
```

## Cannabinoid

```{r}
Manno_RNA %>% rownames() %>%  grep('Cnr[[:digit:]].*',., value=TRUE)

Manno_RNA_Cnr<- Manno_RNA[c('Cnr1', 'Cnr2'),] %>% 
  as.matrix() %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column('Cell_ID')

Manno_RNA_Cnr_merged <- Manno_metadata %>% 
  select(IsCycling,Age,Class,Subclass) %>% 
  rownames_to_column('Cell_ID') %>% 
  inner_join(Manno_RNA_Cnr, by='Cell_ID')

Manno_RNA_Cnr_merged 
```

```{r}
Plot_dotplot_Cnr_by_class <- Manno_RNA_Cnr_merged %>%
  pivot_longer(c(Cnr1, Cnr2), names_to = 'Gene', values_to = 'Expression') %>%  
  mutate_all(~replace(., .==0, NA)) %>%
  group_by(Class, Gene) %>% 
  summarise(Avg=mean(Expression,na.rm=TRUE),
            Percent= 1-(sum(is.na(Expression))/n())) %>% 
  ungroup() %>% 
  arrange(Gene,desc(Avg)) %>% 
  ggplot()+
  geom_point(aes(y=Class, x=Gene, size=Percent, color=Avg))+
  theme_classic()+
  scale_color_gradient(low = "#caf0f8",high = '#03045e') 


ggsave(Plot_dotplot_Cnr_by_class,
       filename='~/Plot_dotplot_Cnr_by_class.pdf',
       device = "pdf",
       height = 6, width =4 , units = "in")
```

```{r}
Plot_dotplot_Cnr_by_class <- Manno_RNA_Cnr_merged %>%
  pivot_longer(c(Cnr1, Cnr2), names_to = 'Gene', values_to = 'Expression') %>%  
  group_by(Class, Gene) %>% 
  summarise(Avg=mean(Expression,na.rm=TRUE),
            Percent= (sum(Expression>0)/n())) %>% 
  filter(Class%in%c('Neural tube','Radial glia','Neuroblast','Neuron')) %>% 
  mutate(Class=factor(Class, levels=c('Neural tube','Radial glia','Neuroblast','Neuron'))) %>% 
  group_by(Gene) %>% 
  mutate(Zscore=(Avg - mean(Avg))/sd(Avg)) %>% 
  ungroup() %>% 
  mutate_all(~replace(., .==0, NA)) %>%
  ggplot()+
  geom_point(aes(x=Class, y=Gene, size=Percent, color=Zscore))+
  theme_classic()+
  scale_color_gradient2(  low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))

ggsave(Plot_dotplot_Cnr_by_class,
       filename='~/Plot_dotplot_Cnr_by_class.pdf',
       device = "pdf",
       height = 3, width =6 , units = "in")
```

```{r}
Plot_dotplot_Cnr_by_class <- Manno_RNA_Cnr_merged %>%
  pivot_longer(c(Cnr1, Cnr2), names_to = 'Gene', values_to = 'Expression') %>%  
  group_by(Class, Gene, Age) %>% 
  summarise(Avg=mean(Expression,na.rm=TRUE),
            Percent= (sum(Expression>0)/n())) %>% 
  filter(Class%in%c('Neural tube','Radial glia','Neuroblast','Neuron')) %>% 
  mutate(Class=factor(Class, levels=c('Neural tube','Radial glia','Neuroblast','Neuron'))) %>% 
  group_by(Gene) %>% 
  mutate(Zscore=(Avg - mean(Avg))/sd(Avg)) %>% 
  ungroup() %>% 
  filter(Gene== 'Cnr1') %>% 
  mutate_all(~replace(., .==0, NA)) %>%
  ggplot()+
  geom_point(aes(x=Class, y=Gene, size=Percent, color=Zscore))+
  theme_classic()+
  scale_color_gradient2(  low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))

ggsave(Plot_dotplot_Cnr_by_class,
       filename='~/Plot_dotplot_Cnr_by_class.pdf',
       device = "pdf",
       height = 3, width =6 , units = "in")
```

```{r}

Manno_mouse <- NormalizeData(object = Manno_mouse, scale.factor = 1e6)
Manno_mouse <- ScaleData(object = Manno_mouse)
Manno_mouse <- FindVariableFeatures(object = Manno_mouse)
Manno_mouse <- RunPCA(object = Manno_mouse, features = VariableFeatures(object = Manno_mouse))
Manno_mouse <- FindNeighbors(Manno_mouse, dims = 1:40)
Manno_mouse <- FindClusters(Manno_mouse, resolution = 0.5)
Manno_mouse <- RunUMAP(Manno_mouse, dims = 1:40)

saveRDS(Manno_mouse, file = '~/Manno_mouse.rds')

Manno_mouse <- readRDS('~/Manno_mouse.rds')

Idents(object = Manno_mouse) <- "Class"

group_colors <- c(Mesenchymal= '#E0E0E0',
                  Other= '#E0E0E0',
                  Neuroepithelial= '#E8E8A2',
                  `Radial Glial`= '#79B5E0',
                  IPC= '#8059B3',
                  Neuronal= '#ED8785') 

Manno_mouse@meta.data <-  Manno_mouse@meta.data %>%
  mutate(color_groups= case_when(Class == 'Neuron' ~ 'Neuronal',
                                 Class %in% c('Neural tube','Neural crest') ~ 'Neuroepithelial',
                                 Class== 'Radial glia' ~ 'Radial Glial',
                                 Class== 'Neuroblast' ~ 'IPC',
                                 TRUE ~ 'Other')) 

pdf(file = '~/feature_manno_fam.pdf')

DimPlot(Manno_mouse, reduction = "umap", 
        group.by = 'Class', label = T, repel = T,  raster = T) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)
  

UMAP_plot_1 <- DimPlot(Manno_mouse, reduction = "umap", 
        group.by = 'color_groups', raster = T) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)+
  ggtitle(element_blank())

  

UMAP_plot_2 <- FeaturePlot(Manno_mouse, reduction = "umap", 
        features = c('Cnr1'), raster = T, order = T)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
  ggtitle(element_blank())



egg::ggarrange(UMAP_plot_1, UMAP_plot_2, ncol=1)

dev.off()
```

```{r}


pdf(file = '~/feature_manno_fam.pdf')

DimPlot(Manno_mouse, reduction = "umap", 
        group.by = 'Class', label = T, repel = T,  raster = T) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)
  

UMAP_plot_1 <- DimPlot(Manno_mouse, reduction = "umap", 
        group.by = 'color_groups', raster = T) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)+
  ggtitle(element_blank())


UMAP_plot_2 <- FeaturePlot(Manno_mouse, reduction = "umap", 
        features = c('Fam72a'), raster = T, order = T)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
  ggtitle(element_blank())

UMAP_plot_3 <- FeaturePlot(Manno_mouse, reduction = "umap", 
        features = c('Srgap2'), raster = T, order = T)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
  ggtitle(element_blank())

egg::ggarrange(UMAP_plot_1, UMAP_plot_2, UMAP_plot_3, ncol=1)

dev.off()
```

### dotplot la manno

```{r}
CNR1_manno <- FetchData(object = Manno_mouse, vars = "Cnr1", layer = "data") %>% rownames_to_column('Cell_Id')

Manno_mouse_meta <-  Manno_mouse@meta.data %>%  rownames_to_column('Cell_Id') %>%  select(Cell_Id,Age,Class)

com_class_order= c('Other','NSC', 'RG', 'IPC', 'NEU')
age_order_manno=c('e7.0', 'e8.0', 'e8.5', 'e9.0', 'e10.0',
  'e11.0', 'e12.0', 'e12.5', 'e13.0','e13.5',
  'e14.0', 'e14.5', 'e15.0','e15.5', 'e16.0',
  'e16.25','e16.5', 'e17.0', 'e17.5','e18.0')

Manno_CNR1_exp_meta <- inner_join(CNR1_manno, Manno_mouse_meta) %>%   
  mutate(com_class= case_when(Class == 'Neuron' ~ 'NEU',
                                 Class %in% c('Neural tube','Neural crest') ~ 'NSC',
                                 Class== 'Radial glia' ~ 'RG',
                                 Class== 'Neuroblast' ~ 'IPC',
                                 TRUE ~ 'Other'),
         com_class= factor(com_class,levels=com_class_order),
         Age=factor(Age, levels=age_order_manno)
         ) 

Manno_dotplot_data <- Manno_CNR1_exp_meta %>%
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(Cnr1,na.rm=TRUE),
            Percent= (sum(Cnr1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Manno')

write_csv(Manno_dotplot_data, '~/manno_dotplot_data_CNR1.csv')  

```




```{r}
Braun_2023_exression <-  read_csv('datasets/human_files/X.csv', col_names = F) %>%
  as.matrix() %>% 
  t() %>% 
  Matrix::Matrix(sparse = T)

gene_metadata <- read_csv('datasets/human_files/var.csv') 
cell_metadata <- read_csv('datasets/human_files/obs.csv')
```

# Eze

```{r}
cell_metadata <- read_tsv('/users/genomics/josema/scdatasets/eze_2021/meta.tsv')
gene_expression_matrix <-  read_tsv('/users/genomics/josema/scdatasets/Eze_2021/exprMatrix.tsv') %>%
  column_to_rownames('gene') %>% 
  as.matrix() %>% 
  Matrix::Matrix(sparse = T)
```

```{r}
library(readxl)
cortex_metadata <- read_excel("/users/genomics/josema/scdatasets/Eze_2021/41593_2020_794_MOESM3_ESM.xlsx", 
    sheet = "5 Cortex Annotations") %>% 
  rename(Cell=`...1`)
```

```{r}
cortex_metadata <- cortex_metadata %>% 
  filter(Cell %in% colnames(gene_expression_matrix)) %>% 
  arrange(factor(Cell, levels=colnames(gene_expression_matrix))) %>% 
  column_to_rownames('Cell')
```

```{r}
Eze <-  CreateSeuratObject(counts = gene_expression_matrix, meta.data = cortex_metadata)
Eze <-  NormalizeData(object = Eze, scale.factor = 1e6)
Eze@assays$RNA$data <-  gene_expression_matrix
Eze <- ScaleData(object = Eze)
Eze <- FindVariableFeatures(object = Eze)
Eze <- RunPCA(object = Eze, features = VariableFeatures(object = Eze))
Eze <- FindNeighbors(Eze, dims = 1:40)
Eze <- FindClusters(Eze, resolution = 0.5)
Eze <- RunUMAP(Eze, dims = 1:40)
```


```{r}
# como no funciona esta mierda porque ya está normalizado, lo hago por ggplot

UMAP_coordinates <-  read_tsv('/users/genomics/josema/datasets/Eze_2021/UMAP.coords.tsv', 
                              col_names = F) %>% 
  setNames(c('Cell', 'UMAP1','UMAP2'))

UMAP_coordinates <-  cbind( UMAP_coordinates,
                            gene_expression_matrix [c('FAM72A','FAM72B','FAM72C','FAM72D',
                                                    'SRGAP2','SRGAP2B','SRGAP2C'),
                                                  UMAP_coordinates$Cell] %>% as.matrix() %>%  t() )
UMAP_coordinates <-  UMAP_coordinates %>% 
  inner_join(cortex_metadata %>%  rownames_to_column('Cell'), by= 'Cell') 


UMAP_coordinates %>%  distinct(`Cell Type`)
```


```{r}

group_colors <- c(Mesenchymal= '#E0E0E0',
                  Other= '#E0E0E0',
                  Neuroepithelial= '#E8E8A2',
                  `Radial Glial`= '#79B5E0',
                  IPC= '#8059B3',
                  Neuronal= '#ED8785')

pdf('~/plot_test.pdf')

UMAP_plot_1 <- UMAP_coordinates %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP1, y= UMAP2, color=CNR1)) +
  geom_point(size=0.1,aes(color=`Cell Type`))+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)


UMAP_Plot_2 <-  UMAP_coordinates %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP1, y= UMAP2, color=CNR1)) +
  geom_point(size=0.1)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)

egg::ggarrange(UMAP_plot_1, UMAP_Plot_2, ncol=1)

dev.off()


```

```{r}
feature_plot_eze <-  function(gen){
  UMAP_coordinates %>% 
  arrange({{gen}}) %>% 
  ggplot( aes(x= UMAP1, y= UMAP2, color={{gen}})) +
  geom_point(size=0.1, show.legend = F)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
    xlab(element_blank())+
    ylab(element_blank())
 }

pdf('~/feature_Eze_fam.pdf', width=9, height=4)

UMAP_coordinates %>% 
  ggplot( aes(x= UMAP1, y= UMAP2)) +
  geom_point(size=0.1,aes(color=`Cell Type`))+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)

FAM72A_plot<-  feature_plot_eze (FAM72A)
FAM72B_plot<-  feature_plot_eze (FAM72B)
FAM72C_plot<-  feature_plot_eze (FAM72C)
FAM72D_plot<-  feature_plot_eze (FAM72D)


SRGAP2_plot<-  feature_plot_eze (SRGAP2)
SRGAP2B_plot<-  feature_plot_eze (SRGAP2B)
SRGAP2C_plot<-  feature_plot_eze (SRGAP2C)



egg::ggarrange(FAM72A_plot, FAM72D_plot, FAM72C_plot, FAM72B_plot,
               SRGAP2_plot, SRGAP2B_plot, SRGAP2B_plot, SRGAP2C_plot, ncol=4)

dev.off()
```

### eze dotplot

```{r}

age_order_eze <- c('1','12','13','14','15','19','20','22')

eze_metadata <- cortex_metadata %>%  
  select( Age, `Cell Type`) %>% 
  rownames_to_column('Cell_Id') %>% 
  mutate(com_class= case_when(`Cell Type` %in% c('Neuronal') ~ 'NEU',
                              `Cell Type`%in% c('Radial Glial') ~ 'RG',
                              `Cell Type`%in% c('IPC') ~ 'IPC',
                              `Cell Type`%in% c('Neuroepithelial') ~ 'NSC',
                              TRUE ~ 'Other'),
         Age=factor(Age, levels=age_order_eze))

eze_dotplot_data <- gene_expression_matrix['CNR1',] %>%
  as.data.frame() %>% setNames('CNR1') %>% 
  rownames_to_column('Cell_Id') %>%
  inner_join(eze_metadata, by= 'Cell_Id') %>% 
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(CNR1,na.rm=TRUE),
            Percent= (sum(CNR1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Eze')

write_csv(eze_dotplot_data,'CNR1/eze_dotplot_data.csv')

```



# NOAK

```{r}
expression <- readRDS("~/merged_scRNA_unfiltered_IDs_faye.RDS")
expression <- UpdateSeuratObject(expression)


group_colors <- c(MG= '#E0E0E0',
                  CR= '#E0E0E0',
                  Mural= '#E0E0E0',
                  IN= '#E0E0E0',
                  `NSC`= '#79B5E0',
                  `NSC_M`= '#79B5E0',
                  IPC= '#8059B3',
                  IPC_M= '#8059B3',
                  PN1= '#ED8785',
                  PN2= '#ED8785',
                  PN3= '#ED8785')


feature_expression_plot <- function(gene){functionFeaturePlot(expression, reduction = "umap",
        features = c(gene), 
        label = F, raster = T, order = T,  pt.size= 2) +
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
  ggtitle(element_blank())
}


pdf(file = '~/feature_noak_fam.pdf')

DimPlot(expression, reduction = "umap", 
        repel = T, label = T, raster = T,  pt.size= 2) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)+
  ggtitle(element_blank())


plot_class <- DimPlot(expression, reduction = "umap", 
                      label = F, raster = T,  pt.size= 2) +
  guides(fill="none", color="none")+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)+
  ggtitle(element_blank())



FAM72D <- FeaturePlot(expression, reduction = "umap",
        features = c('FAM72D'), 
        label = F, raster = T, order = T,  pt.size= 2) +
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)+
  ggtitle(element_blank())

FAM72D <- FeaturePlot(expression, reduction = "umap",
        features = c('FAM72D'), 
        label = F, raster = T, order = T,  pt.size= 2) +
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_clas

egg::ggarrange(plot_class, plot_fam72a, plot_srgap2, ncol=1)


egg::ggarrange(plot_class, plot_cnr1, ncol=1)

dev.off()

```

### Dotplot noak

```{r}
expression <- readRDS("~/merged_scRNA_unfiltered_IDs_faye.RDS")
expression <- UpdateSeuratObject(expression)
CNR1_noak <- FetchData(object = expression, vars = "Cnr1", layer = "data") %>% rownames_to_column('Cell_Id')


metadata<-expression@meta.data
metadata$celltype<-Idents(expression)
metadata$celltype<-gsub("_M","",metadata$celltype)

metadata<-metadata %>% 
  filter(!is.na(pseudotime)) 

age_order_NOAK <-  c(paste('NSC',1:8, sep = '_'), paste('IPC',1:12, sep = '_'),paste('PN1',1:9, sep = '_'),paste('PN2',1:8, sep = '_'),paste('PN3', 1:2, sep = '_'))


grouped_pseudotime <- metadata %>% 
  rownames_to_column('Cell_Id') %>% 
  select(Cell_Id,celltype, pseudotime) %>% 
  group_by(celltype) %>% 
  arrange(pseudotime) %>% 
  mutate(Pseudotime_group=rep(1:ceiling(n()/175),each=175, length.out=n())) %>% 
  group_by(celltype, Pseudotime_group) %>% 
  filter(n()==175) %>% 
  ungroup() %>% 
  mutate(Age= paste(celltype,Pseudotime_group, sep='_'),
        com_class= case_when(celltype %in% c('PN1', 'PN2', 'PN3') ~ 'NEU',
                                 celltype%in% c('NSC', 'NSC_M') ~ 'RG',
                                 celltype%in% c('IPC', 'IPC_M') ~ 'IPC',
                                 TRUE ~ 'Other'),
         Age=factor(Age, levels=age_order_NOAK))


Noak_dotplot_data <- inner_join(CNR1_noak,grouped_pseudotime) %>% 
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(Cnr1,na.rm=TRUE),
            Percent= (sum(Cnr1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Noak')
  
write.csv(Noak_dotplot_data, '~/Noak_dotplot_data')
```


# trevino

```{r}
trevino<-readRDS('~/Trevino/RDS_files/data.RDS')
trevino
saveRDS(trevino_cnr1, '~/Trevino/RDS_files/data_cnr1.RDS')
#trevino <-  NormalizeData(object = trevino, scale.factor = 1e6)
```


```{r}
trevino_cnr1 <-  trevino@meta.data %>% 
  mutate(CNR1 = trevino@assays$RNA$data[c('CNR1'),rownames(trevino@meta.data)],
         )%>% 
  mutate(com_class= case_when(cell.type %in% c('GluN5','GluN4','GluN3','GluN2','GluN7','GluN1','GluN6','GluN8') ~ 'NEU',
                              cell.type%in% c('Early_RG','tRG','Cyc._Prog.','Late_RG') ~ 'RG',
                              cell.type%in% c('nIPC') ~ 'IPC',
                              TRUE ~ 'Other'))

trevino_cnr1 %>% 
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(CNR1,na.rm=TRUE),
            Percent= (sum(CNR1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Trevino') %>% 
  write.csv('~/trevino_dotplot_data')
  

group_colors <- c(
MG = '#E0E0E0',
mGPC = '#E0E0E0',
OPC_Oligo = '#E0E0E0',
GluN5 = '#ED8785',
GluN4 = '#ED8785',
Late_RG = '#79B5E0',
GluN3 = '#ED8785',
Peric. = '#E0E0E0',
SP = '#E0E0E0',
nIPC = '#8059B3',
CGE_IN = '#E0E0E0',
Cyc._Prog. = '#79B5E0',
tRG = '#79B5E0',
MGE_IN = '#E0E0E0',
GluN2 = '#ED8785',
Early_RG = '#79B5E0',
GluN7 = '#ED8785',
EC = '#E0E0E0',
GluN1 = '#ED8785',
GluN6 = '#ED8785',
GluN8 = '#ED8785',
RBC = '#E0E0E0',
VLMC = '#E0E0E0'
 )





pdf('~/trevino_feature.pdf')

a<- trevino_cnr1 %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP_0, y= UMAP_1, color=`cell.type`)) +
  geom_point(size=0.1)+
  theme_classic()+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme(aspect.ratio=1)+
    scale_color_manual(values=group_colors, na.value = 'yellow')

b <- trevino_cnr1 %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP_0, y= UMAP_1, color=CNR1)) +
  geom_point(size=0.1)+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')

egg::ggarrange(a,b,ncol=1)

dev.off()

```

## trevino dotplot




# dotplot fusionado 



# monkey dataset

```{r}
metadata_monkey <- read_csv('/users/genomics/josema/metadata_monkey.csv')
CNR1_monkey <- read_csv('/users/genomics/josema/monkey_CNR1.csv')

metadata_filtered <- metadata_monkey %>%  select(UMAP_1,UMAP_2,orig.ident, observations, subclass)

  

age_oder_monkey <- c('E37','E42','E43','E54','E62','E64','E77','E78','E93','RMB691','RMB683')  

group_colors <- c(`dorsal NSC`='#E8E8A2' ,   
                  `GE NSC`='#E8E8A2',
                  
                  `Early subtypes` =   '#79B5E0',
                  `Patterning centers` = '#79B5E0',
                  
                  enIPC= '#8059B3',
                  inIPC= '#8059B3',

                  CR= '#ED8785',
                  `Excitatory neurons` ='#ED8785',
                  `Inhibitory neurons` ='#ED8785',

                  gIPC ='#E0E0E0',
                  Astro ='#E0E0E0',
                  `OPC-Oligo` ='#E0E0E0',    
                  Immune =  '#E0E0E0',        
                  `RB & Vas` = '#E0E0E0',      
                  Mes ='#E0E0E0')

UMAP_coordinates <- metadata_filtered %>% inner_join(CNR1_monkey) %>%  rename(CNR1='ENSG00000118432')
```


```{r}
pdf('~/plot_monkey.pdf')

UMAP_plot_1 <- UMAP_coordinates %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP_1, y= UMAP_2)) +
  geom_point(size=0.1,aes(color=subclass))+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme_classic()+
  theme(aspect.ratio=1)+
  scale_color_manual(values=group_colors)


UMAP_Plot_2 <-  UMAP_coordinates %>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= UMAP_1, y= UMAP_2, color=CNR1)) +
  geom_point(size=0.1)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)

egg::ggarrange(UMAP_plot_1, UMAP_Plot_2, ncol=1)

dev.off()

mutate( case_when())

```

### monkey dotplot
```{r}
UMAP_coordinates <- UMAP_coordinates %>% 
  mutate(com_class= case_when(subclass %in% c('CR','Excitatory neurons','Inhibitory neurons') ~ 'NEU',
                              subclass%in% c('Early subtypes','Patterning centers') ~ 'RG',
                              subclass%in% c('enIPC','inIPC') ~ 'IPC',
                              subclass%in% c('dorsal NSC','GE NSC') ~ 'NSC',
                              TRUE ~ 'Other'),
         Age=factor(orig.ident, levels=age_oder_monkey))

micali_dotplot_data <- UMAP_coordinates %>% 
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(CNR1,na.rm=TRUE),
            Percent= (sum(CNR1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Micali')

write_csv(micali_dotplot_data,'CNR1/micali_dotplot_data.csv')
```


# dotplot general

```{r}
Noak_dotplot_data <-  read_csv('/home/josema/Desktop/CNR1/Noak_dotplot_data.csv')
Manno_dotplot_data <- read_csv('/home/josema/Desktop/CNR1/manno_dotplot_data_CNR1.csv')
Eze_dotplot_data <- read_csv('/home/josema/Desktop/CNR1/eze_dotplot_data.csv')
Micali_dotplot_data <- read_csv('/home/josema/Desktop/CNR1/micali_dotplot_data.csv')
trevino_dotplot_data<- read_csv('/home/josema/Desktop/CNR1/trevino_dotplot_data.csv')
age_order_NOAK <-  c(paste('NSC',1:8, sep = '_'),
                     paste('IPC',1:12, sep = '_'),
                     paste('PN1',1:9, sep = '_'),
                     paste('PN2',1:8, sep = '_'),
                     paste('PN3', 1:2, sep = '_'))

age_order_eze <- c('1','12','13','14','15','19','20','22')

age_order_manno=c('e7.0', 'e8.0', 'e8.5', 'e9.0', 'e10.0',
                  'e11.0', 'e12.0', 'e12.5', 'e13.0','e13.5',
                  'e14.0', 'e14.5', 'e15.0','e15.5', 'e16.0',
                  'e16.25','e16.5', 'e17.0', 'e17.5','e18.0')

age_order_trevino <-  c('pcw16','pcw20','pcw21','pcw24')

age_oder_monkey <- c('E37','E42','E43','E54','E62','E64',
                     'E77','E78','E93','RMB691','RMB683')  

age_order_bella <- c('E10','E11','E12','E13','E14','E15','E16','E17','E18_S1','E18_S3','P1','P1_S1','P4')
age_order_braun <- paste0(c('5','5.5','6','6.6','6.7','6.9','7','7.5','8','8.1','8.5','9.2','9.5','10','11.5','12','13','14'),'_W')

dibella_dotplot %>% distinct(Age)
global_order <- c(age_order_NOAK,age_order_eze,age_order_manno,age_oder_monkey,age_order_bella,age_order_braun,age_order_trevino)

```


```{r}
merged_dotplot <- rbind(Noak_dotplot_data,
                        Eze_dotplot_data,
                        Manno_dotplot_data,
                        Micali_dotplot_data,
                        dibella_dotplot,
                        braun_dotplot,
                        trevino_dotplot_data[,2:7]) %>% 
  filter(!is.na(Age)) %>% 
  filter(com_class!='Other') %>% 
  mutate(com_class=factor(com_class, levels= c('NSC', 'RG', 'IPC', 'NEU')),
         Age=factor(Age,levels=global_order),
         datasets=dataset) %>% 
  group_by(datasets) %>% 
  mutate(max_0=(Avg- 0) /(max(Avg)-0))
```


```{r}
merged_dotplot_plot <- merged_dotplot %>% 
  ggplot()+
  geom_tile(aes(y=dataset, x=Age, fill=max_0),
             color='gray80')+
  theme_classic()+
  scale_fill_gradient(low = "#caf0f8",high = '#03045e')+
  facet_grid(datasets~dataset+com_class, switch = 'y', scales = 'free', space='free')+
  theme( axis.text.x = element_text( angle = 90,vjust=0.5,hjust = 1),
         legend.position = 'bottom',
         strip.placement = "outside",
         strip.background.x=element_rect(color = NA,
                                        fill=NA), 
        strip.background.y=element_blank(),
        strip.text.y.left = element_blank(),
        strip.text.x.top = element_text(face = 'bold', color = 'gray30'),
        axis.title.y=element_blank(),
        panel.border = element_rect(color = "gray50", fill = NA),
                  panel.grid.major.x  = element_blank() )+
    guides(size=guide_legend(nrow=1, byrow=TRUE,title.position="top", title.hjust = 0.5),
           fill = guide_colourbar(title.position="top", title.hjust = 0.5))+
    scale_size_continuous(range = c(0,2))

ggsave(merged_dotplot_plot,
       filename='/home/josema/Desktop/Braun/merged_dotplot.pdf',
       device = "pdf",
       height = 3.5, width =20, units = "in") 

library(ggplot2)
```


```{r}
plot_function_merged <- function(dataset){
merged_tile <- dataset %>% 
  ggplot()+
  geom_tile(aes(y='CNR1', x=Age, fill=min_max),
             color='gray80')+
  theme_classic()+
  scale_fill_gradient(low = "#caf0f8",high = '#03045e')+
  facet_grid(~dataset+com_class, switch = 'y', scales = 'free', space='free' )+
  theme( axis.text.x = element_text( angle = 45,vjust=1,hjust = 1),
         legend.position = 'bottom',
         strip.placement = "outside",
         strip.background.x=element_rect(color = 'white',
                                        fill='gray40'), 
        strip.background.y=element_blank(),
        strip.text.y.left = element_blank(),
        strip.text.x.top = element_blank(),
        axis.title.y=element_blank(),
        panel.border = element_rect(color = "gray80", fill = NA),
                  panel.grid.major.x  = element_blank() )+
    guides(size=guide_legend(nrow=1, byrow=TRUE,title.position="top", title.hjust = 0.5),
           fill = guide_colourbar(title.position="top", title.hjust = 0.5))


merged_percent <- dataset %>% 
  ggplot()+
  geom_col(aes(y=Percent, x=Age),
           fill='gray80',
           color='gray80')+
  theme_classic()+
  facet_grid(~dataset+com_class, switch = 'y', scales = 'free', space='free' )+
  theme( legend.position = 'bottom',
         strip.placement = "outside",
         strip.background.x=element_rect(color = NA,
                                        fill=NA), 
        strip.background.y=element_blank(),
        strip.text.y.left = element_blank(),
        strip.text.x.top = element_text(face = 'bold', color = 'gray30'),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        panel.border = element_rect(color = NA , fill = NA),
                  panel.grid.major.x  = element_blank() )+
    guides(size=guide_legend(nrow=1, byrow=TRUE,title.position="top", title.hjust = 0.5),
           fill = guide_colourbar(title.position="top", title.hjust = 0.5))

return( egg::ggarrange( merged_percent, merged_tile, ncol=1, heights = c(1.75,1)))
}

pdf('/home/josema/Desktop/Braun/common_tile.pdf', width=30, height=2.8)
plot_function_merged(merged_dotplot)
dev.off()


pdf('CNR1/noak_tile.pdf', width=10, height=3)
plot_function_merged(Noak_dotplot_data)
dev.off()

pdf('CNR1/eze_tile.pdf', width=10, height=2.5)
plot_function_merged(Eze_dotplot_data)
dev.off()

pdf('CNR1/manno_tile.pdf', width=10, height=2.8)
plot_function_merged(Manno_dotplot_data)
dev.off()

pdf('CNR1/micali_tile.pdf', width=10, height=3)
plot_function_merged(Micali_dotplot_data)
dev.off()


ggsave(merged_percent,
       filename='CNR1/merged_tile.pdf',
       device = "pdf",
       height = 7, width =10, units = "in") 


```




# Di bella dataset

```{r}
dibella_expression <- Matrix::readMM('/home/josema/Downloads/gene_sorted-matrix.mtx')
dibella_Metadata <- read_tsv('/home/josema/Downloads/metaData_scDevSC.txt') %>% slice(-1)
dibella_genes <-  read_tsv('/home/josema/Downloads/genes(1).tsv',col_names = F)

colnames(dibella_expression) <- dibella_Metadata %>% pull(NAME)
rownames(dibella_expression) <- dibella_genes %>% pull(X1)

dibella_expression_CNR1 <- dibella_expression['Cnr1',] %>%  as.data.frame() %>%  set_names('CNR1')
dibella_UMAP <- read_tsv('/home/josema/Downloads/cluster_scDevSC.merged.umap.txt')%>% slice(-1)

dibella_merged <- dibella_Metadata %>%  
  cbind(dibella_expression_CNR1) %>% 
  inner_join(dibella_UMAP,by='NAME') %>% 
  mutate(com_class= case_when(New_cellType %in% c('Cajal Retzius cells','CThPN',
                                       'DL CPN','DL_CPN_1','DL_CPN_2',
                                       'Immature neurons','Interneurons',
                                       'Layer 4','Layer 6b','Migrating neurons',
                                       'SCPN','UL CPN','VLMC') ~ 'NEU',
                              New_cellType%in% c('Apical progenitors') ~ 'RG',
                              New_cellType%in% c('Intermediate progenitors') ~ 'IPC',
                              TRUE ~ 'Other'),
         X=as.numeric(X),
         Y=as.numeric(Y))

dibella_dotplot <- dibella_merged %>%  
  mutate(Age=biosample_id) %>% 
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(CNR1,na.rm=TRUE),
            Percent= (sum(CNR1>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='di Bella')
```


```{r}
group_colors <- c(Other= '#E0E0E0',
                  RG= '#79B5E0',
                  IPC= '#8059B3',
                  NEU= '#ED8785')

pdf('~/Downloads/plot_test_dibella.pdf')
 
UMAP_plot_1 <- dibella_merged %>% 
  ggplot( aes(x= (X), y= (Y), color=com_class)) +
  geom_point(size=0.1)+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme_classic()+
  theme(aspect.ratio=1)+
    scale_color_manual(values=group_colors)


UMAP_Plot_2 <-  dibella_merged%>% 
  arrange(CNR1) %>% 
  ggplot( aes(x= (X), y= (Y), color=CNR1)) +
  geom_point(size=0.1)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)

egg::ggarrange(UMAP_plot_1, UMAP_Plot_2, ncol=1)

dev.off()
```

# Braun

```{r}
braun_expression <- read_csv('/home/josema/Desktop/human_CNR1.csv') 
braun_Metadata <- read_csv('/home/josema/Downloads/obs.csv')

braun_ID <- read_csv('/home/josema/Desktop/Braun/CellID.csv')[,2]
braun_tsne <- read_csv('/home/josema/Desktop/Braun/Embedding.csv') [,2:3]
braun_age <- read_csv('/home/josema/Desktop/Braun/Age.csv')[,2]
braun_merged_CNR1<-  cbind(braun_Metadata,braun_expression) 

braun_merged_CNR1 <- cbind(braun_ID, braun_tsne, braun_age) %>% setNames(c('CellID', 'tsne1', 'tsne2', 'Age')) %>% 
  inner_join(braun_merged_CNR1, by='CellID')

braun_merged_CNR1 <- braun_merged_CNR1 %>% 
  mutate(com_class= case_when(CellClass %in% c("b'Neuron'") ~ 'NEU',
                              CellClass%in% c("b'Radial glia'") ~ 'RG',
                              CellClass%in% c("b'Neuronal IPC'"	,"b'Neuroblast'"	) ~ 'IPC',
                              TRUE ~ 'Other'))

group_colors <- c(Other= '#E0E0E0',
                  RG= '#79B5E0',
                  IPC= '#8059B3',
                  NEU= '#ED8785')


pdf('~/Downloads/plot_test_braun.pdf')
 
UMAP_plot_1 <- braun_merged_CNR1 %>% 
    arrange(ENSG00000118432) %>% 
  ggplot( aes(x= (tsne1), y= (tsne2), color=com_class)) +
  geom_point(size=0.1)+
  guides(color = guide_legend(override.aes=list(shape = 15, size = 5)))+
  theme_classic()+
  theme(aspect.ratio=1)+
    scale_color_manual(values=group_colors)


UMAP_Plot_2 <-  braun_merged_CNR1%>% 
  arrange(ENSG00000118432) %>% 
  ggplot( aes(x= (tsne1), y= (tsne2), color=ENSG00000118432)) +
  geom_point(size=0.1)+
  scale_color_gradient(low = "#caf0f8",high = '#03045e')+
  theme_classic()+
  theme(aspect.ratio=1)

egg::ggarrange(UMAP_plot_1, UMAP_Plot_2, ncol=1)

dev.off()

```

## braun dotplot

```{r}

braun_dotplot <- braun_merged_CNR1 %>%  
  group_by(com_class,Age) %>% 
  summarise(Avg=mean(ENSG00000118432,na.rm=TRUE),
            Percent= (sum(ENSG00000118432>0)/n())) %>% 
  ungroup() %>% 
  mutate(min_max=(Avg- min(Avg)) /(max(Avg)-min(Avg))) %>% 
  mutate(dataset='Braun',
         Age=paste0(as.factor(Age),'_W')) 
```


