---
title: "Trevino_analisis"
output: html_document
date: "2024-03-01"
---

Open the libraries

```{r libraries}
library(tidyverse)
library(readr)
library(ggupset)
library(Seurat)
```

# Create SEURAT object

In first place we load the different data files. We have 2 different data classes: The main data files and the metadata. Metadata is divided in two different files: gene metadata and cell metadata.

```{r 'read raw data'}
##### read counts 
count_data<- read_tsv('/projects_ng/SC_DATA/Trevino2019/reannotated/counts.tsv')
count_data<- count_data %>% column_to_rownames(var="...1")

##### read metadata
metadata <- read_tsv('/projects_ng/SC_DATA/Trevino2019/reannotated/meta_cells.tsv')
metadata<- metadata %>% column_to_rownames(var="...1")

genes_metadata <- read_tsv('/projects_ng/SC_DATA/Trevino2019/reannotated/meta_genes.tsv')
genes_metadata<- genes_metadata %>% column_to_rownames(var="...1")
```

Now we do merge the different files creating a seurat object, using `CreateSeuratObject.`

**WARNING**: In python, both data files (count data and metadata) have the same structure (columns= gene names, rows= Cell), but `CreateSeuratObject` uses different data sstructures for counts ((columns= Cell, rows= Genes),) and meta.data (same as python).

```{r seurat object}
# create seurat object
data <- CreateSeuratObject(counts=t(as.matrix(count_data)),
                          # transpose of the count matrix because
                          # we need to set the cell names as columns
                          meta.data=metadata)

# add UMAP data to seurat object 

dimensions <- as.matrix(data@meta.data[,c("UMAP_0","UMAP_1")])
colnames(dimensions) <- c("UMAP_1","UMAP_2")

data[['UMAP']] <- CreateDimReducObject(embeddings = dimensions,
                                       key = "UMAP_",
                                       global = T,
                                       assay = "RNA")
```

Ploteamos el UMAP

```{r UMAP plot}
#Create an UMAP plot of our results 
plot <- Embeddings(data, reduction = 'UMAP') %>%
  cbind(data['reannotated type']) %>% 
  as.data.frame() %>% 
  ggplot(., aes(x=UMAP_1, y=UMAP_2, color=reanotated))+
  geom_point() 

ggsave(plot,
       filename = "/users/genomics/josema/Trevino/umap.pdf",
       device = "pdf",
       height = 5, width = 6, units = "in")
```

# Data analysis

```{r 'data analysis'}

#Normalize data
data <- NormalizeData(data,
                      normalization.method = "LogNormalize",
                      scale.factor = 10000)
#Variable  features (top 10)
data <- FindVariableFeatures(data,
                             nfeatures = 2000)
top10 <- head(VariableFeatures(data), 10)

#plot the variables
VariableFeaturePlot(data) %>% 
  #plot variables names
  LabelPoints(plot = .,
              points = top10,
              repel = TRUE) %>% 
  # save the plot
  ggsave(filename = "/users/genomics/josema/Trevino/feature.pdf",
         device = "pdf",
         height = 5, width = 6, units = "in")

#Scale data
data <- ScaleData(data, 
                  features = rownames(data))
# PCA
data <- RunPCA(data,
               features = VariableFeatures(object = data))
# Neighbours and clusters
data <- FindNeighbors(data, dims = 1:30, reduction = "pca")
data <- FindClusters(data, resolution = 2, cluster.name = "unintegrated_clusters")
```

Now we save the data to create a **checkpoint** for future analysis

```{r checkpoint}
#saveRDS(data, '~/Trevino/RDS_files/data.RDS')
data<-readRDS('~/Trevino/RDS_files/data.RDS')
```

# Filtering and processing data

Now that we have the expression data of all our cell lines genes, we are going to set the genes that are interesting for our analysis. We used the gene list from Xabier's review, based on the paper "The Human Transcription Factors" (Lambert et al 2018)

```{r 'filtering bHLHs'}
# Load dataset of genes that are interesting and extract gene Symbols
table_bhlh<- read_csv('~/Trevino/factors_classification.csv') 
bhlh_symbol <- table_bhlh %>% pull(Symbol) 

# For future analysis we are going to filter only the genes that are highly related with neurogenesis (filtering only neuroG, neuroD, ID and HES)
bhlh_filtered_symbol <- bhlh_symbol %>%
  grep('NEUROD.*|NEUROG.*|TCF.*|ID.*|HES.*',
       .,
       value = T) 

```

# Select bHLH genes

```{r}
# Extract a dataframe with the expression of all  bHLH genes
dataframe_bHLH <- LayerData(data, assay = "RNA", layer = "data") %>% 
  as.data.frame() %>%
  filter(rownames(.)%in%bhlh_symbol) %>% 
  t() %>% 
  as.data.frame()
```

# UpSet plots

In this step we re going to determine for every cell line if a gene is expressed or not. To determine that we could use different thresholds, but as a preliminary analysis **we are going to** **consider that every value different to 0 means expression of a gene**.

To use ggUpset and create an upset plot using ggplot, we have to create a data frame that contains a column with a list of all genes that are expressed on each cell.

```{r 'cargar datos'}
####### load de los datos atajos ######
upset_imput_grouped_by_type <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput_grouped_by_type.RDS')
upset_imput_filtered <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput_filtered.RDS')
upset_imput <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput.RDS')
```

```{r upset plot}
upset_imput <-  dataframe_bHLH%>% 
  rownames_to_column('ID') %>% 
  rowwise() %>%
  mutate(bHLHs = list(names(.)[which(c_across(-ID) != 0)])) %>% 
  select(ID,bHLHs)

reannotated <- data$`reannotated type`%>% 
  as.data.frame() %>% 
  magrittr::set_colnames('reannotated') %>% 
  rownames_to_column('ID')

#Add a column with the annotation of every cell line 
upset_imput <- inner_join(upset_imput,
                          reannotated,
                          by='ID')
```

```{r upset plot}
upset_imput_filtered <-  dataframe_bHLH%>% 
  select(bhlh_filtered_symbol) %>% 
  rownames_to_column('ID') %>% 
  rowwise() %>%
  mutate(bHLHs = list(names(.)[which(c_across(-ID) != 0)])) %>% 
  select(ID,bHLHs)
  
upset_imput_filtered <- inner_join(upset_imput_filtered,
                          reannotated,
                          by='ID')
```

Create groups of transcription factors

```{r upset plot}
# Upset by groups of TF
upset_imput_grouped_by_type<- upset_imput_filtered %>% 
  mutate(NEUROG=ifelse(any(grepl('NEUROG.*',bHLHs)), 'yes','no'),
         NEUROD=ifelse(any(grepl('NEUROD.*',bHLHs)), 'yes','no'),
         TCF=ifelse(any(grepl('TCF.*',bHLHs)), 'yes','no'),
         ID_prot=ifelse(any(grepl('ID.*',bHLHs)), 'yes','no'),
         HES=ifelse(any(grepl('HES.*',bHLHs)), 'yes','no')) %>%
  pivot_longer(cols=-c(ID,bHLHs,reannotated),names_to = 'Genes', values_to = 'Expressed') %>% 
  filter(Expressed=='yes') %>%
  group_by(ID,bHLHs,reannotated) %>% 
  summarize(families=list(Genes))
  

# Extraemos un listado de todos los tipos de anotaciones distintos que hay
reannotation_types <- upset_imput_grouped_by_type %>%
  filter(reannotated!='NA') %>% 
  pull(reannotated) %>%
  unique()
```

upset plot

```{r upset plot}
# Hacemos un upset general en el que se vean todos, quitando los NA
upset_plot_filtered_all <- upset_imput_grouped_by_type %>% 
  filter(reannotated !='NA') %>% 
  ggplot(aes(x = families)) +
  geom_bar(aes(fill=reannotated), position="fill")+
  scale_x_upset()+
  theme_classic()+
  scale_y_continuous(expand = c(0, 0))


ggsave(plot = upset_plot_filtered_all,
       filename ="/users/genomics/josema/Trevino/all_upset.pdf",
       device = "pdf",
       height = 6, width =10 , units = "in")
```

```{r upset plot}
#hacemos un bucle para hacer un upset de cada subtipo celular
for (type in reannotation_types) {
  upset_plot_filtered_group <- upset_imput_grouped_by_type %>% 
    filter(reannotated==type) %>% 
    ggplot(aes(x = families)) +
    ggtitle(type)+
    geom_bar(fill='lightcoral') +
    scale_x_upset()+
    theme_classic()+
    scale_y_continuous(expand = c(0, 0))
  
  
  ggsave(plot = upset_plot_filtered_group,
         filename =str_c("/users/genomics/josema/Trevino/", type, "_upset.pdf"),
         device = "pdf",
         height = 6, width =10 , units = "in")
}
```

```{r upset plot}
upset_imput_grouped_by_type$families <- sapply(upset_imput_grouped_by_type$families[!!lengths(upset_imput_grouped_by_type$families)], toString)

pseudotime_plot<- upset_imput_grouped_by_type %>% 
  mutate(pseudotime=plyr::round_any(velocity_pseudotime*10,1,f=floor)) %>% 
  group_by(families, pseudotime) %>% 
  summarise(n=n()) %>% 
  mutate(families=strsplit(families, ',')) %>% 
  ggplot(aes(y=pseudotime, x=families))+
  geom_point(aes(size=n))+
  coord_flip()+
  ggupset::scale_x_upset()+
  theme_classic()

ggsave(plot = pseudotime_plot,
       filename ='~/pseudotime_plot.pdf',
       device = "pdf",
       height = 6, width =10 , units = "in")
```

```{r upset plot}
imput_plot_upset <- upset_imput_grouped_by_type %>% 
  group_by(reannotated) %>%
  mutate(decile = ntile(velocity_pseudotime, 10)) %>% 
  group_by(reannotated,decile,families) %>% 
  summarise(n=n()) %>% 
  group_by(reannotated,decile) %>% 
  mutate(percent=(n/sum(n)),
         families=strsplit(families, ','))

```

```{r ''}
{for (type in reannotation_types) {}
  pseudotime_plot<- imput_plot_upset%>% 
    filter(reannotated==type) %>% 
    ggplot(aes(y=decile, x=families))+
    ggtitle(type)+
    geom_point(aes(size=percent), color='lightcoral')+
    ggupset::scale_x_upset()+
    theme_classic()
  
  ggsave(plot = pseudotime_plot,
         filename =str_c("/users/genomics/josema/Trevino/upsets_psudotime/",
                         type,
                         "_pseudotime_upset.pdf"),
         device = "pdf",
         height = 6, width =6 , units = "in")
}


```

# How to properly set a threshold of expression

As I found in this response in [Stack Exange](https://bioinformatics.stackexchange.com/questions/5046/find-a-cutoff-value-for-genes-that-are-expressed-in-single-cell-rna-seq):

*"Much of the Seurat workflow incorporates normalized, scaled, and mean-centered expression values.*

*Raw gene expression counts for a cell are normalized by total expression, multiplied by 10,000, and log transformed. Expression level for each gene is scaled by dividing the centered gene expression levels by their standard deviations, resulting in z-scores. It's not an ideal solution, but I find a threshold of 1 (in the realm of z-scores) to makesense. A z-score of 1 is one standard deviation from the mean, or what may be interpreted as baseline expression."*

**So I'll consider the threshold of expression when a z-score value is higher than 1**

```{r}
data<-readRDS('~/Trevino/RDS_files/data.RDS')

upset_imput_grouped_by_type <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput_grouped_by_type.RDS')
upset_imput_filtered <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput_filtered.RDS')
upset_imput <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput.RDS')
```

# reannotation

which annotation systen is the right one? there are multiple columns that are related with the annotation system. If we represent how data changes across multiple classifications using an alluvial plot, we obtain this graphic

```{r}
library(ggalluvial)
plot_classification <- metadata %>% 
  select(`original type`,
         `final type`,
         `was reannot`) %>% 
  filter(`was reannot`=='TRUE') %>% 
  rownames_to_column('cell') %>% 
  count(`original type`, `final type`) %>% 
  ggplot(aes(axis1=`original type`,
             axis2=`final type`,
             y= n)) +
  geom_alluvium(aes(fill = `final type`), show.legend=F) +
  geom_stratum() +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("original type", "`final type`"),
                   expand = c(0.15, 0.05)) +
  theme_void()

ggsave(plot=plot_classification,
       filename='~/Trevino/plot_classification.pdf', 
       device='pdf',height = 10, width =10 , units = "in" )  



```

# Pseudotime distribution across cell types

```{r}


pseudotime_statistics_bygroups <- metadata %>% 
  select(`final type` , velocity_pseudotime) %>% 
  group_by(`final type`) %>%
  filter(`final type`!='NA') %>% 
  summarise(median=median(velocity_pseudotime, na.rm=T),
            mean=mean(velocity_pseudotime, na.rm=T),
            min=min(velocity_pseudotime, na.rm=T)) %>% 
  arrange(min)

pseudotime_statistics_bygroups%>% 
  mutate(across(c(median, mean), ~ rank(.))) %>%
  write_csv('~/Trevino/pseudotime_statistics_bygroups')
  print(n=Inf)
  
   
pseudotime_order_bygroups <- pseudotime_statistics_bygroups %>% 
pull(`final type`)

plot_pseudotime_all <- metadata %>% 
  select(`final type`, `was reannot`, velocity_pseudotime) %>% 
  mutate(`final type`= factor(`final type`,
                              levels=pseudotime_order_bygroups)) %>% 
  filter(`final type`!='NA') %>% 
  ggplot(aes(x=`final type`, y=velocity_pseudotime)) + 
  geom_violin(aes(fill=`final type`),)+
  geom_point(aes(color=`final type`),
             size=0.1,
             alpha = 0.3, 
             position=position_jitter(height=0, width=.4))+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  guides(color = 'none', fill='none')+
  coord_cartesian(ylim = c(0, 1))
  
  

ggsave(plot_pseudotime_all,
       filename='~/Trevino/pseusotime_groups_plot.pdf',
       device = "pdf",
       height = 6, width =6 , units = "in")
```

# Check if data is pre-filtered

Using count data, I'm going to plot the distribution of:

-   How many genes are expressed in every cell. (To determine if they've previously filtered cells that don't have a minimum number of expressed genes)

-   How many cells express every gene. (To determine if they've previously filtered genes that are not expressed in a minimum number of cells)

```{r}

cell_expression_distribution <- rowSums(count_data != 0)
gene_expression_distribution <- colSums(count_data != 0)

values <- c(cell_expression_distribution, gene_expression_distribution)


plot_dataset <- data.frame(values=values,
                           class=c(rep('n_genes_expressed',
                                       length(cell_expression_distribution)),
                                   rep('expressed_in_n_cells',
                                       length(gene_expression_distribution))
                                   )
                           )
# Print the dataframe with the count

plot_filter_analysis <- plot_dataset %>% 
  ggplot(aes(x=class, y=values))+
  geom_violin(aes(fill=class))+
  geom_point(aes(color=class),
             size=1,
             alpha = 0.05,
             position=position_jitter(height=0, width=.4))+
  theme_bw()+
  theme(axis.text.x = element_blank())+
  guides(color = 'none', fill='none')+
  facet_wrap(~class,scales = 'free')

ggsave(plot_filter_analysis,
       filename='~/Trevino/plot_filter_analysis.pdf',
       device = "pdf",
       height = 6, width =6 , units = "in")


```

# ver el desarrollo del pseudotime a lo largo de los diferentes grupos

```{r}
library(RColorBrewer)
UMAP_velocity_plot <- metadata %>% 
  ggplot(aes(x=UMAP_0, y=UMAP_1, color=velocity_pseudotime))+
  geom_point(size=1)+
  theme_bw()+
  theme(aspect.ratio=1)+
  scale_color_distiller(palette = 'Blues',  trans = "reverse")+
  labs(color='Pseudotime')

ggsave(UMAP_velocity_plot,
       filename='~/Trevino/plot_umap_velocity.pdf',
       device = "pdf",
       height = 6, width =7 , units = "in")

UMAP_types_plot <- metadata %>% 
  ggplot(aes(x=UMAP_0, y=UMAP_1, color=`final type`))+
  geom_point(size=1)+
  theme_bw()+
  theme(aspect.ratio=1)+
  labs(color='Pseudotime')

ggsave(UMAP_types_plot,
       filename='~/Trevino/plot_umap_types.pdf',
       device = "pdf",
       height = 6, width =7 , units = "in")
```

# UpSet pero teniendo en cuenta el orden de los tipos celulares según el pseudotime

```{r}
generate_combinations_dataset <- function(original_dataset) {
  
  # Function to generate all possible combinations
  generate_combinations <- function(elements) {
    all_combinations <- unlist(lapply(1:length(elements), function(x) combn(elements, x, simplify = FALSE)), recursive = FALSE)
    all_combinations <- lapply(all_combinations, sort)
    all_combinations <- unique(all_combinations)
    all_combinations
  }
  
  # Function to check if a row has all elements in a combination
  check_combination <- function(row_elements, combination) {
    all(combination %in% row_elements)
  }
  
  # Extract elements
  elements <- unlist(original_dataset$List)
  
  # Generate all possible combinations
  all_combinations <- generate_combinations(elements)
  
  # Create new dataset
  new_dataset <- original_dataset %>%
    mutate(across(all_combinations, ~map_int(.x, ~check_combination(elements, .))))
  
  return(new_dataset)
}

# Example usage:
original_dataset <- tibble(
  ID = c(1, 2, 3),
  List = list(c("A", "B"), c("B", "C"), c("A", "B", "C"))
)

result_dataset <- generate_combinations_dataset(original_dataset)
print(result_dataset)

```

# Co-binding evolution plot

```{r}
Co_binding <- upset_imput_grouped_by_type %>% 
  ungroup() %>% 
  mutate(decile = ntile(velocity_pseudotime, 100)) %>% 
  group_by(decile,families) %>% 
  summarise(n=n()) %>% 
  group_by(decile) %>% 
  mutate(percent=(n/sum(n))) %>% 
  mutate(families=stringi::stri_replace_all_regex(families,
                                                  pattern=c(', ','HES','NEUROD',
                                                          'NEUROG', 'ID_prot','TCF'),
                                                  replacement=c('+','H','ND',
                                                                'NG','ID','TCF'),
                                                  vectorize=FALSE))



for(group in c('H','ND','NG','ID','TCF')){
co_binfing_plot <- Co_binding %>%
  mutate(families=factor(families, levels= Co_binding %>% 
                                              arrange(str_length(families))%>% 
                                              pull(families) %>% 
                                              unique()
                         ) 
         ) %>% 
  ggplot(aes(x=decile,y=percent, color=families))+
    geom_line()+
    # ggrepel::geom_label_repel(aes(label = families, color=families),
    #              data = Co_binding %>% filter(decile == 100) %>%
    #                filter(grepl('H',families)),
    #              nudge_x = 0.35,
    #              max.overlaps=Inf,
    #              show.legend = FALSE,
    #              fill= alpha("white", 0.7))+
  gghighlight::gghighlight(grepl(group,families),
              use_direct_label=FALSE)+
    facet_wrap(~families)+
    # geom_line(data= Co_binding, aes(group=families),color='gray90')+
    theme_classic()+
  guides(fill="none", color="none")+
  coord_cartesian(expand = F)+
  theme(plot.margin = margin(0.5,0.5,0.5,0.5, "cm"))
  

  ggsave(plot = co_binfing_plot,
         filename =str_c("/users/genomics/josema/Trevino/co_binfing_plot", group, ".pdf"),
         device = "pdf",
         height = 6, width =6 , units = "in")
}
```

```{r}
Co_binding_reanalysis <- upset_imput_grouped_by_type %>% 
  ungroup() %>% 
  mutate(decile = ntile(velocity_pseudotime, 100)) %>%
  group_by(decile,families) %>% 
  summarise(n=n()) %>% 
  group_by(families) %>% 
  mutate(percent=(n/sum(n))) %>% 
  mutate(families=stringi::stri_replace_all_regex(families,
                                                  pattern=c(', ','HES','NEUROD',
                                                          'NEUROG', 'ID_prot','TCF'),
                                                  replacement=c('+','H','ND',
                                                                'NG','ID','TCF'),
                                                  vectorize=FALSE))



for(group in c('H','ND','NG','ID','TCF')){
co_binfing_plot <- Co_binding_reanalysis %>%
  mutate(families=factor(families, levels= Co_binding_reanalysis %>% 
                                              arrange(str_length(families))%>% 
                                              pull(families) %>% 
                                              unique()
                         ) 
         ) %>% 
  filter(grepl(group,families)) %>% 
  ggplot(aes(x=decile,y=percent, color=families))+
    geom_col()+
    facet_wrap(~families, scales = 'free_y')+
    # geom_line(data= Co_binding, aes(group=families),color='gray90')+
  theme_void()+
  guides(fill="none", color="none")+
  coord_cartesian(expand = F)+
  theme(plot.margin = margin(0.5,0.5,0.5,0.5, "cm"))
  

  ggsave(plot = co_binfing_plot,
         filename =str_c("/users/genomics/josema/Trevino/co_binding_plot_2", group, ".pdf"),
         device = "pdf",
         height = 6, width =6 , units = "in")
}
```

```{r}
generate_combinations <- function(input_list) {
  output <- list()
  n <- length(input_list)
  
  # Include individual elements
  output <- append(output, input_list)
  
  # Include combinations of elements
  for (i in 2:n) {
    combs <- combn(input_list, i, simplify = FALSE)
    output <- append(output, combs)
  }
  
  return(output)
}

# Example usage
input_list <- c('A', 'B', 'C')
combinations <- generate_combinations(input_list)
print(combinations)
```

```{r}
library(tidyverse)

generate_combinations_dataset <- function(original_dataset) {
  # Function to generate all possible combinations
  generate_combinations <- function(elements) {
    all_combinations <- unlist(lapply(1:length(elements), function(x) combn(elements, x, simplify = FALSE)), recursive = FALSE)
    all_combinations <- lapply(all_combinations, sort)
    all_combinations <- unique(all_combinations)
    all_combinations
  }
  
  # Function to check if a row has all elements in a combination
  check_combination <- function(row_elements, combination) {
    all(combination %in% row_elements)
  }
  
  # Extract elements
  elements <- original_dataset$List
  
  # Generate all possible combinations
  all_combinations <- generate_combinations(unlist(elements))
  
  # Create new dataset
  new_dataset <- original_dataset %>%
    mutate(across(all_combinations, ~map_int(.x, ~check_combination(elements, .))))
  
  return(new_dataset)
}

# Example usage:
original_dataset <- tibble(
  ID = c(1, 2, 3),
  List = list(c("A", "B"), c("B", "C"), c("A", "B", "C"))
)

result_dataset <- generate_combinations_dataset(original_dataset)
print(result_dataset)

```

```{r}
library(tidyverse)

create_combinations_dataset <- function(data) {
  # Function to generate combinations
  generate_combinations <- function(elements) {
    all_combinations <- unlist(lapply(1:length(elements), function(x) combn(elements, x, simplify = FALSE)), recursive = FALSE)
    all_combinations <- lapply(all_combinations, sort)
    all_combinations <- unique(all_combinations)
    all_combinations
  }
  
  # Function to check if a row has all elements in a combination
  check_combination <- function(row_elements, combination) {
    all(combination %in% row_elements)
  }
  
  # Generate all possible combinations
  all_combinations <- data %>% 
    pull(List) %>% 
    unlist() %>% 
    unique() %>% 
    generate_combinations()
  
  # Create new dataset
  new_dataset <- data %>% 
    mutate(across(all_combinations, ~map_int(.x, ~check_combination(List, .))))
  
  return(new_dataset)
}

# Sample data
input_data <- tibble(
  ID = c(1, 2, 3),
  List = list(c("A", "B"), c("B", "C"), c("A", "B", "C"))
)

# Generate new dataset
output_data <- create_combinations_dataset(input_data)

# Print the output dataset
print(output_data)
```

# Tabla transformada en local

```{r}
#Excribimos una tabla en tsv para poder leerla en python y hacer el análsis en python

upset_imput_grouped_by_type %>% 
  ungroup() %>% 
  select(ID, families) %>% 
  write_tsv('/users/genomics/josema/upset_imput_grouped_by_type.tsv')
```

```{python}
# load the packages
import pandas as pd
from itertools import combinations

# create a function that creates all the combinations in a table
def create_combinations_dataset(data, sep):
    # Convert the string of elements into a list
    data['families'] = data['families'].str.split(sep)
    
    # Get unique elements
    unique_elements = set().union(*data['families'])
    
    # Generate all possible combinations
    all_combinations = ['+'.join(comb) for i in range(len(unique_elements)+1) for comb in combinations(unique_elements, i) if comb]
    
    # Create a new DataFrame
    new_data = pd.DataFrame()
    new_data['ID'] = data['ID']
    
    # Fill new DataFrame with 0 or 1 based on combinations
    for comb in all_combinations:
        new_data[comb] = data['families'].apply(lambda x: int(all(elem in x for elem in comb.split('+'))))
    
    return new_data

# Sample data
input_data = pd.read_table('/users/genomics/josema/upset_imput_grouped_by_type.tsv')

# Generate new dataset
output_data = create_combinations_dataset(input_data, sep=', ')

# Print the output dataset
output_data.to_csv('/users/genomics/josema/prueba.tsv', sep='\t')

```

# volvemos a R

```{r}
# cargamos la tabla que he creado en 
recursive_data <- read.delim('/users/genomics/josema/prueba.tsv', header=T, row.names = 1) %>%  as.tibble()

# upset_imput_grouped_by_type <- readRDS('/users/genomics/josema/Trevino/RDS_files/upset_imput_grouped_by_type.RDS') 
#   
# upset_imput_grouped_by_type$families <- sapply(upset_imput_grouped_by_type$families[!!lengths(upset_imput_grouped_by_type$families)], toString)

metadata %>% select(Cell.ID,velocity_pseudotime,`final type`)
  
```

# 
